Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-22 surfaces-write-up/surface-write-up.tex
edits to lemma 4.2

\begin{lem}
\label{lem:hirz-generators}
$R_D$ is generated in degree at most
\[
	\sum_{i \in \Te(D)} \gcd(\bida_i, \bidb_i) \ell_i + \sum_{\substack
	{i \in \Tp(D) \\ j\in \Tm(D)}} (\bida_i \bidb_j - \bida_j \bidb_i)
	\ell	_{i,j}.
\]
\end{lem}

\begin{proof}
%Let $(\bida_i, \bidb_i)$ be the bi-degrees of $f_i$. 
Suppose $\gamma
\in (R_D)_d$. \todo{Aaron: Why is $\gamma$ defined? It doesn't seem to be used anywhere.}
 Then 
\[
	g = \sum_{i = 1}^n {f_i}^{c_i}
\]

\noindent
where each $c_i \ge - \alpha_i c_i$, $\sum_{i=1}^n c_i \bida_i = 0$,
and $\sum_{i=1}^n c_i \bidb_i = 0$. We can any view $\sigma$ inside the
lattice 
\begin{align*}
	\Sigma = \{ (d, c_1, \ldots, c_n) \in \bz_{\geq 0}^{n+1} : c_i \ge - d
\alpha_i \text{ for all }i \text{ and }\sum_{i=1}^n c_i \bida_i = \sum_{i =
1}^n c_i \bidb_i = 0 \}.	
\end{align*}

@@

Let $\Delta^{n-1}$ be the simplex with vertices $e_1, \ldots e_n$.
We can intersect $\Delta^{n-1}$ with the hyperplane $H$ given by
$\sum_{i=1}^n \bida_i X_i = 0$  to get a subspace $S \subseteq \Delta
^{n-1}$. The extremal rays of $S$ must lie on the edges of $\Delta
^{n - 1}$ \todo{Peter: Do we need to cite/argue this, or is it clear?}.
Thus, the extremal rays of $S$ are given by the $e_i$'s
contained in $H$ and the intersection points $e_{i, j}$ of $H$
and the edges $\overline{e_i e_j}$ of the simplex between
distinct $e_i, e_j$ not contained in $H$.
\todo{more detail above}

We can calculate elements of $R_D$ whose
corresponding lattice points lie on these extremal rays:

For $i \in \Te(D)$, define \todo{below section needs to be simplified, maybe
by introducing new variables}
\[
	\epsilon_i := (\prod_{k \ne i} f_k^{-\alpha_k \ell_i \gcd(\bida_i, \bidb_i)})
	(f_i^{\ell_i \frac{\gcd(\bida_i, \bidb_i)}{\bida_i + \bidb_i}\sum_{k \ne i}
	\alpha_k (\bida_k + \bidb_k)}) \in H^0(\ell_i \gcd(\bida_i, \bidb_i) D).
\]

\noindent
For $i \in \Tp(D)$ and $j \in \Tm(D)$, define
\[
	\epsilon_{i, j} := \left(\prod_{k \ne i, j} f_k^{-\alpha_k \ell_{i,j} (\bida_i \bidb_
	j - \bida_j \bidb_i)}) {f_i}^{s_1} {f_j}^{s_2} \in H^0(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j \bidb_i) D \right)
\]

@@

The set of $e_i$ and $e_{i,j}$ given above satisfies the conditions of
Proposition ~\ref{prop:cone-generation}, $R_D$ is generated in 
degrees less than the sum of the degrees of the $e_i$ and $e_{i,j}$, which is
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\todo{Aaron: Do you use $\sigma$ later? We are usually reserving $\sigma$ for a
lattice point. Perhaps something like $\delta$ would be more appropriate if you
want to refer to it later. Also, you should number if if you want to refer to
it later. In general, it's bad style to define things in the proofs. This could
have been defined prior to the theorem statement as well, which would simplify
the statement. We should also then clarify that $\sigma$ depends on $D$.}

>>>>

\begin{lem}
\label{lem:hirz-generators}
$R_D$ is generated in degree at most
\[
	\sigma = \sum_{i \in \Te(D)} \gcd(\bida_i, \bidb_i) \ell_i + \sum_{\substack
	{i \in \Tp(D) \\ j\in \Tm(D)}} (\bida_i \bidb_j - \bida_j \bidb_i)
	\ell	_{i,j}.
\]
\end{lem}

\begin{proof}
%Let $(\bida_i, \bidb_i)$ be the bi-degrees of $f_i$. 
Suppose $g
\in (R_D)_d$.
 Then 
\[
	g = \sum_{i = 1}^n {f_i}^{c_i}
\]

\noindent
for some $c_i \ge - \alpha_i d$ such that $\sum_{i=1}^n c_i \bida_i = 0$
and $\sum_{i=1}^n c_i \bidb_i = 0$. We can view $\sigma$ inside the
lattice 
\begin{align*}
	\Sigma = \{ (d, c_1, \ldots, c_n) \in \bz_{\geq 0}^{n+1} : c_i \ge - d
\alpha_i \text{ for all }i \text{ and }\sum_{i=1}^n c_i \bida_i = \sum_{i =
1}^n c_i \bidb_i = 0 \}.	
\end{align*}

@@

Let $\Delta^{n-1}$ be the simplex with vertices $e_1, \ldots e_n$.
We can intersect $\Delta^{n-1}$ with the hyperplane $H$ given by
$\sum_{i=1}^n \bida_i X_i = 0$  to get a subspace $S \subseteq \Delta
^{n-1}$.  Since $\Sigma \subseteq S$, if each extremal rays of $S$ contains an element of the lattice $\Sigma$, then those representatives in $\Sigma$ are in fact the extremal rays of $\Sigma$.

The extremal rays of $S$ must lie on the edges of $\Delta
^{n - 1}$, \todo{Peter: Do we need to cite/argue this, or is it clear?}
and are given by the $e_i$'s
contained in $H$ and the intersection points $e_{i, j}$ of $H$
together with the edges $\overline{e_i e_j}$ of the simplex between
distinct $e_i, e_j$ not contained in $H$.
\todo{more detail above;  Peter: resolved?}
The following are elements of $R_D$ whose corresponding lattice point lies on each of the above extremal rays:

For $i \in \Te(D)$, define \todo{below section needs to be simplified, maybe
by introducing new variables}
\begin{equation}\label{eqn:epsilon_i}
	\epsilon_i := (\prod_{k \ne i} f_k^{-\alpha_k \ell_i \gcd(\bida_i, \bidb_i)})
	(f_i^{\ell_i \frac{\gcd(\bida_i, \bidb_i)}{\bida_i + \bidb_i}\sum_{k \ne i}
	\alpha_k (\bida_k + \bidb_k)}) \in H^0(\ell_i \gcd(\bida_i, \bidb_i) D).
\end{equation}

\noindent
For $i \in \Tp(D)$ and $j \in \Tm(D)$, define
\begin{equation}\label{eqn:epsilon_i,j}
	\epsilon_{i, j} := \left(\prod_{k \ne i, j} f_k^{-\alpha_k \ell_{i,j} (\bida_i \bidb_
	j - \bida_j \bidb_i)} {f_i}^{s_1} {f_j}^{s_2}\right) \in H^0(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j \bidb_i) D
\end{equation}

@@

The set of $e_i$'s and $e_{i,j}$'s corresponding to the $\epsilon_i$'s and $\epsilon_{i,j}$'s given above satisfies the conditions of
Proposition ~\ref{prop:cone-generation}, implying that $R_D$ is generated in 
degrees less than the sum of the degrees of the $e_i$ and $e_{i,j}$, which is
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\todo{Aaron: Do you use $\sigma$ later? We are usually reserving $\sigma$ for a
lattice point. Perhaps something like $\delta$ would be more appropriate if you
want to refer to it later. Also, you should number if if you want to refer to
it later. In general, it's bad style to define things in the proofs. This could
have been defined prior to the theorem statement as well, which would simplify
the statement. We should also then clarify that $\sigma$ depends on $D$.}
